Water, High-Altitude Condensates, and Possible Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}

### 2018-02-26 arxiv_v3/ms.tex
arxiv v3

\title{Water, High-Altitude Condensates, and \added{Possible} Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}

@@

\begin{abstract}
	The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar. The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern. \added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.}  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height. A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations. \added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude.} Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

WASP-107b also has a relatively low equilibrium temperature compared to most other exoplanets that are amenable to atmosphere characterization (780 K, assuming zero albedo).  This results in a distinct atmospheric chemistry compared to other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.  Spectral features from both water and methane are accessible with current observing facilities -- \added{both of these molecules have strong absorption bands in the wavelength range covered by \emph{HST}/WFC3.  Measuring the shape and amplitude of these spectral features enables a constraint on the abundance of water, the dominant oxygen-bearing molecule, and methane, the dominant carbon-bearing molecule, providing a bounded estimate of the carbon-to-oxygen ratio (C/O).} Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage, \added{and have mainly resulted in upper limits} \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.

@@

\subsection{Spectroscopic Light Curve Fits}
We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.63 $\mu$m, shown in Figure\,\ref{fig:lc}. We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics have the same morphology at all wavelengths \citep{stevenson14c, kreidberg14a}. For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.  One advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction. The amplitude of the feature has no detectable wavelength dependence at the level of precision of our data.  As for the white light curve, we fixed some of the transit parameters on the best fit values from \cite{dai17}. We did not put put priors on the transit parameters because varying them simply shifts the spectrum up or down by a constant value. Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties. In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions \added{for a star with $T_\mathrm{eff} = 4300 K$, $\log g = 4.5$, and $[\mathrm{Fe/H}] = 0.$} (coefficients listed in Table\,\ref{tab:tspec}.  The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.  We refer to this fitting approach as ``Method A".

@@

We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum. In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve. \added{This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent within $0.5\sigma$ on average with method A over the bandpass.} We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit. The results from these tests are nearly identical to our nominal transmission spectrum, except with a small constant offset due to the uncorrected star-spot crossing feature. This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}). 
	\added{We also repeated the calculations from Method A, but fit} for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values. \added{This approach, which we dub ``Method C", also yields consistent results with Method A (the relative transit depths agree to better than 0.2 sigma on average). The transit depths and fitted limb darkening coefficients are provide in Table\,\ref{tab:tspec}.} The fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths.  
	\added{To test that uncertainties on the stellar parameters do not significantly bias the limb darkening coefficients, we calculated PHOENIX model limb coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1$). The uncertainty in $T_\mathrm{eff}$ has the largest effect on the estimated limb darkening coefficients. We recalculated the transmission spectrum with limb darkening models that varied $T_\mathrm{eff}$ from 4310 to 4550 K. Over this temperature range, the change in limb darkening shifts the average transit depth by less than 20 ppm, which is below the level of noise in the data.  The relative change in transit depth is smaller still ($< 1$ ppm).}

@@

\added{\subsubsection{Impact of Unocculted Star Spots}
	Unocculted star spots may influence the measured transmission spectrum \citep{mccullough14, zellem17}. To estimate the impact of unocculted spots on our data,  we first estimated the star spot properties based on the observed photometric variability in the Kepler band pass, which is 0.3\% \citep{dai17,mocnik17}. This variability can be reproduced with a single spot with covering fraction of 1\% and temperature 300 K lower than the stellar photosphere.  We calculated the effect of unocculted spots with these properties using equation (1) from \cite{mccullough14}, and predict that they will introduce a positive slope in the transmission spectrum of 8 ppm over the WFC3 bandpass, well below our measurement uncertainties. We note that cool stars may have persistent spot coverage that does not produce any photometric variability \citep{rackham18}. However, even if there is persistent spot coverage of 5\%, the spots produce a slope in the WFC3 bandpass of 50 ppm, which is still below our uncertainties.}

@@

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the main retrieval results were unchanged.  We also explored cases without the $30\times$ solar metallicity prior upper limit. Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.}

@@

\subsubsection{Methane Abundance}
\added{If methane is present in the atmosphere, it has a strong spectral feature centered at $1.4\,\mu$m. This feature overlaps the water feature, but has a wider shape. Our data are precise enough to distinguish this subtle difference in feature shape, so } to explore whether methane depletion is the underlying cause of the inferred low C/O, we ran two additional retrievals.  In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$).  We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium. The $3\,\sigma$ upper limit on methane abundance is $1.4\times10^{-3}$. \added{This value is in tension with expectations for a solar abundance pattern: for $3.7\times$ solar metallicity, $650$ K, and $5\times10^{-4}$ bar (the median retrieved photosphere properties), the expected methane abundance is $1.2\times10^{-3}\times$ solar (for solar C/O).} Based on these tests, we conclude that the atmosphere of WASP-107b is \added{likely} depleted in methane relative to expectations for a solar abundance pattern. By contrast, the water abundance from the free retrieval ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.

%we calculate the thermochemical equilibrium methane abundance for ($1.2\time10^{-3}$ at 650 K, $5\times10^4$ bar, for a solar C/O and $3.7\,\times$ solar metallicity composition).}  We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval; \added{however, we note that the expected equilibrium abundance is sensitive to the true atmosphere metallicity. For a $1\times$ solar composition, the expected methane abundance drops by $1/3$. Similarly, the expected abundances are sensitive to the temperature and pressure in the photosphere. For example, at slightly higher temperatures and higher altitudes (1 mbar, 800 K), the expected methane abundance drops to $2.4\times10^{-5}$.}

\subsection{Condensate Properties}
In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.  We modeled clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a solar composition atmosphere. We varied the cloud sedimentation efficiency from 3 to 0.3 (normal to highly lofted, small particulate clouds).  \added{Only the model with the most lofted clouds ($f_\mathrm{sed} = 0.3$) produces sufficiently small amplitude features to match the observed spectrum.  This result is in agreement with findings for other small planets, such as the low sedimentation efficiency inferred for GJ 1214b \citep[$f_\mathrm{sed} < 0.1$][]{morley15}.}  
%cite Peter Gao's paper?

@@

We also modeled an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere. We predicted the abundance of hydrocarbon haze precursors from previously published $50\times$ solar photochemical models for GJ 436b \citep{line11, morley17}. \added{We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%. We found that the larger particles ($0.3\,\mu$m) with moderate haze production efficiency (3\%) match the shape and amplitude of the spectral features well. More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.}

@@

\begin{itemize}
\item{\emph{The upper limit on atmospheric metallicity from interior structure modeling is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment trend, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.  Compared to results for other exoplanets of similar mass such as GJ 436b, which has a high metallicity ($>100\times$ solar) and HAT-P-26b, which is metal-poor compared to the Solar System trend, WASP-107b adds to the evidence that exoplanets exhibit a greater diversity of compositions than is present in the Solar System \citep{morley17, wakeford17}.}
\item{\emph{The methane abundance is \added{likely} depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.  Another possibility is that the planet has a hot interior effective temperature ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}. Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}. Further observations of the transmission spectrum over a broader wavelength range will refine the C/O estimate and help differentiate between these two scenarios.} 
\item{\emph{Optically thick condensates are present at high altitudes} ($0.1 - 3$ mbar). The amplitude of the water absorption feature in the transmission spectrum is less than a third that expected for a clear atmosphere.  \added{We find that physically-motivated cloud and haze formation models can satisfactorily reproduce the observed feature amplitude. Either lofted clouds with low sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or $0.3\,\mu$m haze particles with moderate haze production efficiency (3\%) match the spectrum.} Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
\end{itemize}

>>>>

\title{Water, High-Altitude Condensates, and Possible Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}

@@

\begin{abstract}
	The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar. The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern. The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height. A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations. We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude. Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

WASP-107b also has a relatively low equilibrium temperature compared to most other exoplanets that are amenable to atmosphere characterization (780 K, assuming zero albedo).  This results in a distinct atmospheric chemistry compared to other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.  Spectral features from both water and methane are accessible with current observing facilities -- both of these molecules have strong absorption bands in the wavelength range covered by \emph{HST}/WFC3.  Measuring the shape and amplitude of these spectral features enables a constraint on the abundance of water, the dominant oxygen-bearing molecule, and methane, the dominant carbon-bearing molecule, providing a bounded estimate of the carbon-to-oxygen ratio (C/O). Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage, and have mainly resulted in upper limits \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.

@@

\subsection{Spectroscopic Light Curve Fits}
We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.63 $\mu$m, shown in Figure\,\ref{fig:lc}. We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics have the same morphology at all wavelengths \citep{stevenson14c, kreidberg14a}. For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.  One advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction. The amplitude of the feature has no detectable wavelength dependence at the level of precision of our data.  As for the white light curve, we fixed some of the transit parameters on the best fit values from \cite{dai17}. We did not put put priors on the transit parameters because varying them simply shifts the spectrum up or down by a constant value. Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties. In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions for a star with $T_\mathrm{eff} = 4300 K$, $\log g = 4.5$, and $[\mathrm{Fe/H}] = 0.$ (coefficients listed in Table\,\ref{tab:tspec}.  The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.  We refer to this fitting approach as ``Method A".

@@

We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum. In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve. This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent within $0.5\sigma$ on average with method A over the bandpass. We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit. The results from these tests are nearly identical to our nominal transmission spectrum, except with a small constant offset due to the uncorrected star-spot crossing feature. This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}).

We also repeated the calculations from Method A, but fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values. This approach, which we dub ``Method C", also yields consistent results with Method A (the relative transit depths agree to better than 0.2 sigma on average). The transit depths and fitted limb darkening coefficients are provide in Table\,\ref{tab:tspec}. The fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths.

To test that uncertainties on the stellar parameters do not significantly bias the limb darkening coefficients, we calculated PHOENIX model limb coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1$). The uncertainty in $T_\mathrm{eff}$ has the largest effect on the estimated limb darkening coefficients. We recalculated the transmission spectrum with limb darkening models that varied $T_\mathrm{eff}$ from 4310 to 4550 K. Over this temperature range, the change in limb darkening shifts the average transit depth by less than 20 ppm, which is below the level of noise in the data.  The relative change in transit depth is smaller still ($< 1$ ppm).

@@

\subsubsection{Impact of Unocculted Star Spots}
Unocculted star spots may influence the measured transmission spectrum \citep{mccullough14, zellem17}. To estimate the impact of unocculted spots on our data,  we first estimated the star spot properties based on the observed photometric variability in the Kepler band pass, which is 0.3\% \citep{dai17,mocnik17}. This variability can be reproduced with a single spot with covering fraction of 1\% and temperature 300 K lower than the stellar photosphere.  We calculated the effect of unocculted spots with these properties using equation (1) from \cite{mccullough14}, and predict that they will introduce a positive slope in the transmission spectrum of 8 ppm over the WFC3 bandpass, well below our measurement uncertainties. We note that cool stars may have persistent spot coverage that does not produce any photometric variability \citep{rackham18}. However, even if there is persistent spot coverage of 5\%, the spots produce a slope in the WFC3 bandpass of 50 ppm, which is still below our uncertainties.

@@

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the main retrieval results were unchanged.  We also explored cases without the $30\times$ solar metallicity prior upper limit. Without this prior, the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.

@@

\subsubsection{Methane Abundance}
If methane is present in the atmosphere, it has a strong spectral feature centered at $1.4\,\mu$m. This feature overlaps the water feature, but has a wider shape. Our data are precise enough to distinguish this subtle difference in feature shape, so to explore whether methane depletion is the underlying cause of the inferred low C/O, we ran two additional retrievals.  In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$).  We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium. The $3\,\sigma$ upper limit on methane abundance is $1.4\times10^{-3}$. This value is in tension with expectations for a solar abundance pattern: for $3.7\times$ solar metallicity, $650$ K, and $5\times10^{-4}$ bar (the median retrieved photosphere properties), the expected methane abundance is $1.2\times10^{-3}\times$ solar (for solar C/O). Based on these tests, we conclude that the atmosphere of WASP-107b is likely depleted in methane relative to expectations for a solar abundance pattern. By contrast, the water abundance from the free retrieval ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.

\subsection{Condensate Properties}
In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.  We modeled clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a solar composition atmosphere. We varied the cloud sedimentation efficiency from 3 to 0.3 (normal to highly lofted, small particulate clouds).  Only the model with the most lofted clouds ($f_\mathrm{sed} = 0.3$) produces sufficiently small amplitude features to match the observed spectrum.  This result is in agreement with findings for other small planets, such as the low sedimentation efficiency inferred for GJ 1214b \citep[$f_\mathrm{sed} < 0.1$][]{morley15}.  
%cite Peter Gao's paper?

@@

We also modeled an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere. We predicted the abundance of hydrocarbon haze precursors from previously published $50\times$ solar photochemical models for GJ 436b \citep{line11, morley17}. We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%. We found that the larger particles ($0.3\,\mu$m) with moderate haze production efficiency (3\%) match the shape and amplitude of the spectral features well. More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.

@@

\begin{itemize}
\item{\emph{The upper limit on atmospheric metallicity from interior structure modeling is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment trend, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.  Compared to results for other exoplanets of similar mass such as GJ 436b, which has a high metallicity ($>100\times$ solar) and HAT-P-26b, which is metal-poor compared to the Solar System trend, WASP-107b adds to the evidence that exoplanets exhibit a greater diversity of compositions than is present in the Solar System \citep{morley17, wakeford17}.}
\item{\emph{The methane abundance is likely depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.  Another possibility is that the planet has a hot interior effective temperature ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}. Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}. Further observations of the transmission spectrum over a broader wavelength range will refine the C/O estimate and help differentiate between these two scenarios.} 
\item{\emph{Optically thick condensates are present at high altitudes} ($0.1 - 3$ mbar). The amplitude of the water absorption feature in the transmission spectrum is less than a third that expected for a clear atmosphere.  We find that physically-motivated cloud and haze formation models can satisfactorily reproduce the observed feature amplitude. Either lofted clouds with low sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or $0.3\,\mu$m haze particles with moderate haze production efficiency (3\%) match the spectrum. Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
\end{itemize}
